Equalizer 2.0 --\\Convergence of a Parallel Rendering Framework

### 2017-12-07 paperIEEE.tex
Review related work
git-svn-id: https://dev.ifi.uzh.ch/vmml/svn/Equalizer16@69 db7c848b-7755-4df2-b763-8796797afa52

Highlight integration of many features into a general framework vs one-off
research implementation of one algorithm. Also point out the various and very
diverse examples in Section 7, highlighting the variety of application and
implementation
scenarios that can benefit from \textsf{Equalizer}.

@@

In~\cite{EMP:09}, we presented the initial \textsf{Equalizer} parallel
rendering framework,
which summarized the work in parallel rendering up to 2009. An extensive
Programming and User Guide provides in-depth documentation on using and
programming \textsf{Equalizer}~\cite{Eilemann:13}. In the following we assume
these two
publications and their references as a starting point, and focus on the related
work published since then.

The concept of transparent OpenGL interception popularized by \textsf{WireGL}
and \textsf{Chromium}~\cite{HHNFAKK:02} has received little attention since
2009. While some commercial implementations such as \textsf{TechViz} and
\textsf{MechDyne Conduit} continue to exist, on the research side only
\textsf{ClusterGL}~\cite{NHM:11} has been presented. \textsf{ClusterGL} employs
the same approach as \textsf{Chromium}, but delivers a significantly faster
implementation of transparent OpenGL interception and distribution for parallel
rendering. \textsf{CGLX}~\cite{DK:11} tries to bring parallel execution
transparently to OpenGL applications, by emulating the GLUT API and intercepting
certain OpenGL calls. In contrast to frameworks like \textsf{Chromium} and
\textsf{ClusterGL} which distribute OpenGL calls, \textsf{CGLX} follows the
distributed application approach. This works transparently for trivial
applications, but quickly requires the application developer to address the
complexities of a distributed application, when mutable application state needs
to be synchronized across processes. For realistic applications, writing
parallel applications remains the only viable approach for scalable parallel
rendering, as shown by the success of \textsf{Paraview}, \textsf{Visit} and
\textsf{Equalizer}-based applications.

On the other hand, software for driving and interacting with tiled display walls
has received significant attention, including \textsf{Sage}~\cite{Sage} and
\textsf{Sage~2}~\cite{Sage2} in particular. \textsf{Sage} was built entirely
around the concept of a shared framebuffer where all content windows are
separate applications using pixel streaming but is no longer actively supported.
\textsf{Sage 2} is a complete, browser-centric reimplementation where each
application is a web application distributed across browser instances.
\textsf{DisplayCluster}~\cite{DC}, and its continuation
\textsf{Tide}~\cite{tide}, also implement the shared framebuffer concept of
\textsf{Sage}, but provide a few native content applications integrated into the
display servers. These solutions implement a scalable display environment and
are a target display platform for scalable 3D graphics applications.

@@

Furthermore, various applications and frameworks have used \textsf{Equalizer}
for new research in visualization. On the application side, \textsf{RTT
Deltagen}, \textsf{Bino}, \textsf{Livre} and \textsf{RTNeuron}~\cite{HBBES:13}
are the most mature examples and are presented in \sref{sApplications}. On the
framework side, \textsf{Omegalib}~\cite{Omegalib}, a framework used in the
Cave2, made significant progress in integrating 2D collaborative workspaces
like \textsf{Sage~2} with 3D immersive content.

visualization for remote sensing data~\cite{LK:09}

\cite{CKP:12} tiled display wall virtual environment

>>>>

\FIXME{Highlight integration of many features into a general framework vs one-off
research implementation of one algorithm. Also point out the various and very
diverse examples in Section 7, highlighting the variety of application and
implementation
scenarios that can benefit from \textsf{Equalizer}.}

@@

In 2009 we presented the initial \textsf{Equalizer} parallel rendering
framework~\cite{EMP:09}, which introduced the architecture of a generic parallel
rendering frameworks and summarized the work in parallel rendering. Since then,
an extensive Programming and User Guide provides in-depth documentation on using
and programming \textsf{Equalizer}~\cite{Eilemann:13}. In the following related
work we assume these two publications and their references as a starting point,
and focus on the new work published since 2009.

The concept of transparent OpenGL interception popularized by \textsf{WireGL}
and \textsf{Chromium}~\cite{HHNFAKK:02} has received little attention since
2009. While some commercial implementations such as \textsf{TechViz} and
\textsf{MechDyne Conduit} continue to exist, on the research side only
\textsf{ClusterGL}~\cite{NHM:11} has been presented. \textsf{ClusterGL} employs
the same approach as \textsf{Chromium}, but delivers a significantly faster
implementation of transparent OpenGL interception and distribution for parallel
rendering. \textsf{CGLX}~\cite{DK:11} tries to bring parallel execution
transparently to OpenGL applications, by emulating the GLUT API and intercepting
certain OpenGL calls. In contrast to frameworks like \textsf{Chromium} and
\textsf{ClusterGL} which distribute OpenGL calls, \textsf{CGLX} follows the
distributed application approach. This works transparently for trivial
applications, but quickly requires the application developer to address the
complexities of a distributed application, when mutable application state needs
to be synchronized across processes. For realistic applications, writing
parallel applications remains the only viable approach for scalable parallel
rendering, as shown by the success of \textsf{Paraview}, \textsf{Visit} and
various \textsf{Equalizer}-based applications.

On the other hand, software for driving and interacting with tiled display walls
has received significant attention, including \textsf{Sage}~\cite{Sage} and
\textsf{Sage~2}~\cite{Sage2} in particular. \textsf{Sage} was built entirely
around the concept of a shared framebuffer where all content windows are
separate applications using pixel streaming. It is no longer actively supported.
\textsf{Sage 2} is a complete, browser-centric reimplementation where each
application is a web application distributed across browser instances.
\textsf{DisplayCluster}~\cite{DC}, and its continuation
\textsf{Tide}~\cite{tide}, also implement the shared framebuffer concept of
\textsf{Sage}, but provide a few native content applications integrated into the
display servers. All these solutions implement a scalable display environment and
are a target display platform for scalable 3D graphics applications.

@@

Furthermore, various applications and frameworks have used \textsf{Equalizer}
for new research in visualization. On the application side, \textsf{RTT
Deltagen}, \textsf{Bino}, \textsf{Livre} and \textsf{RTNeuron}~\cite{HBBES:13}
are the most mature examples and are presented in \sref{sApplications}. On the
framework side, \textsf{Omegalib}~\cite{Omegalib}, a framework used in the
Cave2, made significant progress in integrating 2D collaborative workspaces like
\textsf{Sage~2} with 3D immersive content. Lambers et.al. developed a framework
for visualizing remote sensing data~\cite{LK:09} on large displays and immersive
installations based on Equalizer.
